Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-15 introduction.tex
more typos in section 2

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\xi = \frac{4\alpha\nu}{\pi(2\alpha-1)}$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n \ge 3$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

>>>>

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\xi = \frac{4\alpha\nu}{\pi(2\alpha-1)}$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n \ge \GR{2}$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

### 2018-11-15 proof_overview.tex
more typos in section 2

To analyze expression in the both the finite and infinite model we will make use of the~\emph{Campbell-Mecke} formula. 
For a Poisson point process $\mathcal{P}$ on a measurable space $S$ with intensity $\lambda$ and a measurable non-negative function $h: S^r \rightarrow \mathbb{R}$ we have
\begin{equation}\label{eq:def_Campbell-Mecke}
\begin{split}
& \Exp{\sum_{x_1,\ldots, x_r \in \mathcal{P}}^{\neq} h (x_1, \ldots, x_r,
\mathcal{P} \setminus \{x_1,\ldots, x_r\}) }\\
& = \int_S \cdots \int_S
\Exp{h(x_1,\ldots, x_r,\mathcal{P} \setminus \{x_1,\ldots, x_r\}) \lambda (x_1) \cdots \lambda (x_r) dx_1 \cdots dx_r},
\end{split}
\end{equation}
where the sum ranges over all distinct $r$-tuples of points of $\mathcal{P}$.

Using the Campbell-Mecke formula we can compute the density of the ball $\BallPo{p}$ as
\begin{equation}\label{eq:mu_ball_y_P}
	\mu_{\alpha, \nu}(\BallPo{p}) = \int_0^\infty \int_{-\infty}^\infty \ind{|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}}
	f_{\alpha,\nu}(x^\prime, y^\prime) \, dx^\prime \, dy^\prime 
	= \frac{4\alpha \nu}{(2\alpha - 1)\pi} e^{\frac{y}{2}} := \xi_{\alpha,\nu}e^{\frac{y}{2}},
\end{equation}

We shall use $\rho(p,k)$ and $\rho(y,k)$ (for a point $p=(x,y) \in \mathcal{R}$) to denote the probability mass function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(\BallPo{p})$, i.e.
\[
	\rho(p,k) = \rho(y,k) = \Prob{\Po\left(\xi_{\alpha,\nu} e^{y/2}\right) = k}.
\]
Note that $\rho(y,k)$ denote the probability that a node $(x,y) \in G_{\Pcal}(\alpha,\nu)$ has degree $k$, which due to the translation symmetry of the Poisson Point Process in the $x$ direction depends only on the \emph{height} $y$.

@@

Again, due to symmetry in the $x$ direction we can assume without loss of generality that a randomly sampled point $p$ is $(0,y)$. For any $y \in \R_+$ we define the probability density function on $\Rcal := \R_+ \times \R$ by  
\begin{equation}\label{eq:def_eta_P}
	\eta_y(x^\prime, y^\prime) = \frac{\ind{p^\prime \in \BallPo{(0,y)}} f_{\alpha,\nu}(x^\prime,y^\prime)}{\mu_{\alpha,\nu}(\BallPo{(0,y)})}.
\end{equation}
That is $\eta_y$ is the density $f_{\alpha,\nu}$ restricted to the ball $\BallPo{(0,y)}$, properly normalized. Or in other words, it denotes the probability density of a randomly sampled point in $\BallPo{(0,y)}$, i.e. a randomly sampled neighbor of $(0,y)$.

@@

In Section~\ref{sec:asymptotics_average_clustering_ast_P} we will compute both these quantities and show that $c_\Pcal = c_\infty$ and $c_\Pcal(k) = c_\infty(k)$, where $c_\infty$ and $c_\infty(k)$ are given in, respectively, Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:local_clustering_hyperbolic}

@@

To compute the limit of the adjusted local clustering $c_{\mathcal{P},n}^\ast(k)$ in the finite graph $G_{\mathcal{P},n}(\alpha, \nu)$ we first prove in Section~\ref{sec:concentration_c_P_n} that it is concentrated around its mean $\Exp{c_{\mathcal{P},n}^\ast(k_n)}$.

\begin{proposition}[Concentration local clustering in $G_{\mathcal{P},n}(\alpha, \nu)$]\label{prop:concentration_local_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any non-decreasing sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\mathcal{P},n}^\ast(k_n) - \Exp{c_{\mathcal{P},n}^\ast(k_n)}\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

@@

\begin{proposition}[Transition to infinite limit model]\label{prop:convergence_average_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence tending to infinity and satisfying $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\Pcal(k_n)\right| = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

>>>>

To analyze expressions in the both the finite and infinite model we will make use of the~\emph{Campbell-Mecke} formula. 
For a Poisson point process $\mathcal{P}$ on a measurable space $S$ with intensity $\lambda$ and a measurable non-negative function $h: S^r \rightarrow \mathbb{R}$ we have
\begin{equation}\label{eq:def_Campbell-Mecke}
\begin{split}
& \Exp{\sum_{x_1,\ldots, x_r \in \mathcal{P}}^{\neq} h (x_1, \ldots, x_r,
\mathcal{P} \setminus \{x_1,\ldots, x_r\}) }\\
& = \int_S \cdots \int_S
\Exp{h(x_1,\ldots, x_r,\mathcal{P} \setminus \{x_1,\ldots, x_r\}) \lambda (x_1) \cdots \lambda (x_r) dx_1 \cdots dx_r},
\end{split}
\end{equation}
where the sum ranges over all distinct $r$-tuples of points of $\mathcal{P}$.

Using the Campbell-Mecke formula we can compute the intensity measure of the ball $\BallPo{p}$, i.e. the expected number of points falling into it, as
\begin{equation}\label{eq:mu_ball_y_P}
	\mu_{\alpha, \nu}(\BallPo{p}) = \int_0^\infty \int_{-\infty}^\infty \ind{|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}}
	f_{\alpha,\nu}(x^\prime, y^\prime) \, dx^\prime \, dy^\prime 
	= \frac{4\alpha \nu}{(2\alpha - 1)\pi} e^{\frac{y}{2}} := \xi_{\alpha,\nu}e^{\frac{y}{2}},
\end{equation}

We shall use $\rho(p,k)$ and $\rho(y,k)$ (for a point $p=(x,y) \in \mathcal{R}$) to denote the probability mass function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(\BallPo{p})$, i.e.
\[
	\rho(p,k) = \rho(y,k) = \Prob{\Po\left(\xi_{\alpha,\nu} e^{y/2}\right) = k}.
\]
Note that $\rho(y,k)$ denotes the probability that a vertex $(x,y) \in G_{\Pcal}(\alpha,\nu)$ has degree $k$, which due to the translation symmetry of the Poisson Point Process in the $x$ direction depends only on the \emph{height} $y$.

@@

Again, due to symmetry in the $x$ direction we can assume without loss of generality that a randomly sampled point $p$ is $(0,y)$. For any $y \in \R_+$ we define the probability density function on $\Rcal :=  \R \times \R_+$ by  
\begin{equation}\label{eq:def_eta_P}
	\eta_y(x^\prime, y^\prime) = \frac{\ind{p^\prime \in \BallPo{(0,y)}} f_{\alpha,\nu}(x^\prime,y^\prime)}{\mu_{\alpha,\nu}(\BallPo{(0,y)})}.
\end{equation}
That is $\eta_y$ is the density $f_{\alpha,\nu}$ restricted to the ball $\BallPo{(0,y)}$, properly normalized. Or in other words, it denotes the probability density of a randomly sampled point in $\BallPo{(0,y)}$, i.e. a randomly sampled neighbor of $(0,y)$.

@@

In Section~\ref{sec:asymptotics_average_clustering_ast_P} we will compute both these quantities and show that $c_\Pcal = c_\infty$ and $c_\Pcal(k) = c_\infty(k)$, where $c_\infty$ and $c_\infty(k)$ are given in, respectively, Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:local_clustering_hyperbolic}.

@@

To compute the limit of the modified local clustering function $c_{\mathcal{P},n}^\ast(k)$ in the finite graph $G_{\mathcal{P},n}(\alpha, \nu)$ we first prove in Section~\ref{sec:concentration_c_P_n} that it is concentrated around its expectation $\Exp{c_{\mathcal{P},n}^\ast(k_n)}$.

\begin{proposition}[Concentration local clustering function in $G_{\mathcal{P},n}(\alpha, \nu)$]\label{prop:concentration_local_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any non-decreasing sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\Exp{\left|c_{\mathcal{P},n}^\ast(k_n) - \Exp{c_{\mathcal{P},n}^\ast(k_n)}\right|} = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

@@

\begin{proposition}[Transition to infinite limit model]\label{prop:convergence_average_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence tending to infinity and satisfying $k_n = \smallO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\Pcal(k_n)\right| = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}
\MS{Is it really necessary that the sequence tends to infinity in the thm above?}
